\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities\\ Technical Report with Proofs and Details

### 2018-11-07 fully-abstract-comp/technical-report.tex
Fix the worst typesetting problems in the TR.

\title{\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities\\
  Technical Report with Proofs and Details}
\author{Lau Skorstengaard \and Dominique Devriese \and Lars Birkedal}
\begin{document}
\maketitle
\tableofcontents
Disclaimer: While the proofs in this technical report are done, the text can be lacklustre and from time to time out of date.
We will make the technical report up to date as soon as possible.
Further, the use of coloring in the later parts of this document has not always been done with as much care as it ought to. This means that not all source specific things are colored blue, but it should be evident from the context that they belong to the source language.

@@

\lau{09-07-2018: Fix these definitions' formatting like in paper} 
\[
  \lrol = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
    \begin{array}{l}
      \forall i \leq n \ldotp \\
      \quad \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm[i]{\ta,\stkb} \\\qquad\Rightarrow (\ms_T,\reg_T) \term\\
    \end{array}
\right\}
\]
\[
  \lror = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
    \begin{array}{l}
      \forall i \leq n \ldotp \\ 
      \quad (\ms_T ,\reg_T) \term[i] \\\qquad\Rightarrow \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm{\ta,\stkb}
    \end{array}
\right\}
\]

@@

\[
  \lrrg{\trust}(R)(W) = \left\{ \npair{\stpair{\reg}{\reg}} \middle|
    \begin{array}{l}
      \exists S : (\RegName \setminus (\{\pcreg \} \cup R))\fun \World \ldotp \\
      \quad W = \bigoplus_{r \in (\RegName\setminus (\{\pcreg,\rdata \} \cup R))} S(r) \wedge \\
      \quad \forall r \in \RegName \setminus (\{\pcreg \} \cup R)\ldotp\\
      \qquad\npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrvg{\trust}(S(r))
    \end{array}
            \right\}
\]
We write $\lrr(W)$ to mean $\lrr(\emptyset)(W)$. That is, if we do not need to exclude extra registers, then we simply omit that argument.
\lau{RESOLVED check what conditions need to have a ``square'' i.e. need approximation}
\[
  \lrv(W) =
  \begin{array}[t]{l}
    \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
    \begin{array}{l}
      (\isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T}) \wedge\\
      \exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
      \quad H_\sigma \; \sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
      \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
      \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
      \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
      \quad (\nonLinear{\src{\vsc_S}} \Rightarrow \\
      \qquad \forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))

      % This case was never used (?) so it is not necessary.
      % \wedge \\
      % \quad \npair[n']{\src{\vsc_S'},\src{\vsc_S},\vsc_T',\vsc_T} \in \lrexj(W' \oplus W_o)
    \end{array}
    \right\}\cup\\
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}},\\ \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \endarray \right)} 
    \; \middle| \;
    \begin{array}{l}
      [\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss) \tand\\
      \forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
      \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal (\lrv \circ \xi)
    \end{array}
    \right\} \cup \\
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
    \begin{array}{l}
      \begin{array}{r l l}
        \perm \not\in \{\rx,\rwx\} \wedge\\
        % \perm = \noperm & \Rightarrow & \npair{(\linear,\baddr,\eaddr)} \in \lrp(W) \wedge \\
        \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \stackReadCond{W} \wedge \\
        \perm \in \writeAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \stackWriteCond{W}
      \end{array}
    \end{array}
    \right\} \cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
    \begin{array}{l}
      [b,e] \mathrel{\#} \ta \tand\\
      \begin{array}{r l l }
        % \perm = \noperm & \Rightarrow & \npair{(\lin,\baddr,\eaddr)} \in \lrp(W) \wedge \\
        \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \readCond{\lin,W} \wedge\\
        \perm \in \writeAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \writeCond{\lin,W} \wedge\\
        % we are excluding rwx pointers.
        \perm \neq \rwx \wedge \\
        % \perm = \rwx &\Rightarrow& \array[t]{l}\npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge \\
        %                            \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \endarray\\
        \perm = \rx &\Rightarrow& \array[t]{l}\npair{[\baddr,\eaddr]} \in \execCond{W} \wedge\\
        \npair{[\baddr,\eaddr]} \in \xReadCond{W} \wedge \\
                                  \lin = \normal \\ \endarray
      \end{array}
    \end{array}
    \right\}
  \end{array}
\]

\[
  \lrvtrusted[\square,(\ta,\stkb,\gsigrets,\gsigcloss)](W) =
  \begin{array}[t]{l}
    \lrv(W)\cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}},\\ \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \endarray \right)} 
    \; \middle| \;
    \begin{array}{l}
      \exists r \in \dom(\pwheap) \ldotp \\
      \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)} \wedge\\
      \quad \dom(\code) \subseteq \ta \wedge [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss) \wedge\\
      \quad \sigrets \subseteq \gsigrets \wedge \sigcloss \subseteq \gsigcloss
    \end{array}
    \right\} \cup \\
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\normal),\baddr,\eaddr,\aaddr)},\\ ((\perm,\normal),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
    \begin{array}{l}
      \perm \sqsubseteq \rx \wedge \\
      \gc = (\ta,\stkb,\gsigrets,\gsigcloss)  \wedge \\
      {} [\baddr,\eaddr] \subseteq \ta \wedge\\
      \npair{[\baddr,\eaddr]} \in \xReadCond[\square,\gc]{W} 
    \end{array}
    \right\}
  \end{array}
\]

@@

\begin{lemma}
  \label{lem:code-reg-and-mem-sat}
  If $\pwheap(r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$ and
  $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, then there exists $W''$
  and $W'$ such that $W = W' \oplus W''$ and
  $\npair[n']{(\code,\code)} \in H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; \gc \; \xi^{-1}(W')$ for all $n' < n$
\end{lemma}
\begin{proof}
  By definition of $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, we get that $W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$ and
  $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in
  \lrheap(\pwheap)(W_{\var{heap}})$ for some $\ms_S' \subseteq \ms_S$ and $\ms_{T,\var{heap}}\subseteq \ms_T$.

  By definition of $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in
  \lrheap(\pwheap)(W_{\var{heap}})$, we get an
  $R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg$ and $\exists R_W : \dom(\activeReg{\pwheap}) \fun \World$ such that $W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r)$ and $\npair[n']{R_\ms(r_{\var{code}})} \in  \pwheap(r_{\var{code}}).H \; \xi^{-1}(R_W(r_{\var{code}}))$ for all $n' < n$.

@@

\begin{lemma}[non-linear words are pure]
  \label{lem:non-linear-pure}

  \begin{itemize}
  \item If $\npair{A} \in \readCond{\normal,W}$, then $\npair{A} \in \readCond{}(\normal, \purePart{W} )$.
  \item If $\npair{A} \in \writeCond{\normal,W}$, then $\npair{A} \in \writeCond{}(\normal,\purePart{W})$
  \item If $\npair{A} \in \execCond{W}$, then $\npair{A} \in \execCond{}(\purePart{W})$
  \item If $\npair{A} \in \xReadCond{W}$, then $\npair{A} \in \xReadCond{}(\purePart{W} )$.
  \item If $\npair{(w_1,w_2)} \in \lrv(W)$ and ($\nonLinear{w_1}$ or
    $\nonLinear{w_2}$), then
    $\npair{(w_1,w_2)} \in \lrv(\purePart{W})$.
  \end{itemize}
\end{lemma}
\begin{proof}
  Follows easily by inspecting the definitions of $\lrv$, $\readCond[]{}$, $\addressable{}$, $\writeCond[]{}$, $\execCond[]{}$ and $\xReadCond[]{}$ and using Lemma~\ref{lem:purePart-idempotent}.
\end{proof}

\begin{lemma}[permission-based conditions shrinkable]
  \label{lem:conds-shrinkable}
  \begin{itemize}
  \item If $\npair{A} \in \readCond{\lin,W}$  and $\emptyset \neq A' \subseteq A$, then
    $\npair{A'} \in \readCond{\lin,W}$.
  \item If $\npair{A} \in \writeCond{\lin,W}$  and $\emptyset \neq A' \subseteq A$, then
    $\npair{A'} \in \writeCond{\lin,W}$.
  \item If $\npair{A} \in \xReadCond{W}$  and $A' \subseteq A$, then
    $\npair{A'} \in \xReadCond{W}$.
  \item If $\npair{A} \in \execCond{W}$  and $\emptyset \neq A' \subseteq A$, then
    $\npair{A'} \in \execCond{W}$.
  \end{itemize}
\end{lemma}
\begin{proof}
  Follows easily from the definitions.
  In the case for $\execCond{}$ where $\lin = \linear$, we have a partition of the world into $W = \oplus_{a \in A} W_a(a)$ that we need to convert into a partition $W = \oplus_{a \in A'} W_a'(a)$.
  It suffices construct new $W_a'$ by taking $W_a$ but adding $\oplus_{a\in(A\setminus A')}$ to an $W_a(a)$ for an arbitrary $a \in A'$ to make this work.
\end{proof}

@@

   Take $R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg$ such that
    \begin{itemize}
    \item $R_\ms(r_\addr) = (\msdata|_{\{a\}}, \msdata|_{\{a\}})$.
    \item $R_\ms(r_\var{code}) = (\mspad\uplus\mscode,\mspad\uplus\mscode)$
    \end{itemize}
    Take $R_W' : \dom(\activeReg{\pwheap}) \fun \World$ such that
    \begin{itemize}
    \item $R_W'(r_\addr) = \purePart{W}[\var{heap}.r_{\addr'}\mapsto \stdreg{\{\aaddr\},\gc}{\spatialo}]_{\addr' \in A_\var{own}(\addr)}$ 
    \item $R_W'(r_\var{code}) = \purePart{W}$
    \end{itemize}
    Since then $W = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W'(r)$, we can use Lemma~\ref{lem:oplus-future-distr} to construct an $R_W$ with  $W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r)$ and $R_W(r) \future R_W'(r)$ for all $r$.
    Finally, take $R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal}$ to satisfy
    \begin{itemize}
    \item $R_\var{seal}(r_\addr) = \emptyset$
    \item $R_\var{seal}(r_\var{code}) = \sigrets\uplus \sigcloss$.
    \end{itemize}

@@

  From $\npair{(\var{comp},\var{comp})} \in \lrcomp(W)$, we know that
  $\{(\_ \mapsto c_{\mathrm{main},c}), (\_ \mapsto c_{\mathrm{main},d})\}\subseteq \overline{\var{export}}$ and 
  $\npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W)$ with $\var{comp}_0 = (\mscode, \msdata, \overline{a_{\mathrm{import}}\mapsfrom s_{\mathrm{import}}}, \overline{s_{\mathrm{export}}\mapsto w_{\mathrm{export}}}, \sigrets, \sigcloss,A_\linear)$.
  From this, it follows that for all  $W' \future W$, if
  \begin{equation*}
    \overline{\npair[n']{(w_{\mathrm{import}},w_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}
  \end{equation*}
  and $\msdata' = \msdata{}[\overline{a_{\mathrm{import}} \mapsto w_{\mathrm{import}}}]$
  then
  \begin{equation*}
    \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W')
  \end{equation*}
  and
  \begin{equation*}
    \overline{\npair{(w_{\mathrm{export}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
  \end{equation*}
  Since $\overline{w_{\mathrm{import}}} = \emptyset$, the former holds vacuously, $\msdata' = \msdata$ and the latter two results follow for every $W' \future W$.

>>>>

\title{\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities\\
  Technical Report with Proofs and Details}
\author{Lau Skorstengaard \and Dominique Devriese \and Lars Birkedal}
\begin{document}
\maketitle

This document is a technical report accompanying a paper by the same title and authors, published at POPL 2019.
It contains proofs and details that were omitted from the paper for space and presentation reasons.

\tableofcontents
Disclaimer: While the proofs in this technical report are done, the text can be lacklustre and from time to time out of date.
We will make the technical report up to date as soon as possible.
Further, the use of coloring in the later parts of this document has not always been done with as much care as it ought to. This means that not all source specific things are colored blue, but it should be evident from the context that they belong to the source language.

@@

\lau{09-07-2018: Fix these definitions' formatting like in paper} 
\[
  \lro[\preceq,(\ta,\stkb,\_,\_)] = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
    \begin{array}{l}
      \forall i \leq n \ldotp \\
      \quad \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm[i]{\ta,\stkb} \\\qquad\Rightarrow (\ms_T,\reg_T) \term\\
    \end{array}
\right\}
\]
\[
  \lro[\succeq,(\ta,\stkb,\_,\_)] = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
    \begin{array}{l}
      \forall i \leq n \ldotp \\ 
      \quad (\ms_T ,\reg_T) \term[i] \\\qquad\Rightarrow \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm{\ta,\stkb}
    \end{array}
\right\}
\]

@@

\[
  \lrrg{\trust}(R)(W) = \left\{ \npair{\stpair{\reg}{\reg}} \middle|
    \begin{array}{l}
      \exists S : (\RegName \setminus (\{\pcreg \} \cup R))\fun \World \ldotp \\
      \quad W = \bigoplus_{r \in (\RegName\setminus (\{\pcreg,\rdata \} \cup R))} S(r) \wedge \\
      \quad \forall r \in \RegName \setminus (\{\pcreg \} \cup R)\ldotp\\
      \qquad\npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrvg{\trust}(S(r))
    \end{array}
            \right\}
\]
We write $\lrr(W)$ to mean $\lrr(\emptyset)(W)$. That is, if we do not need to exclude extra registers, then we simply omit that argument.
\lau{RESOLVED check what conditions need to have a ``square'' i.e. need approximation}
\[
  \lrv(W) =
  \begin{array}[t]{l}
    \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
%
    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
    \begin{array}{l}
      (\isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T}) \wedge\\
      \exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
      \quad H_\sigma \; \sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
      \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
      \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
      \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
      \quad (\nonLinear{\src{\vsc_S}} \Rightarrow \\
      \qquad \forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))

      % This case was never used (?) so it is not necessary.
      % \wedge \\
      % \quad \npair[n']{\src{\vsc_S'},\src{\vsc_S},\vsc_T',\vsc_T} \in \lrexj(W' \oplus W_o)
    \end{array}
    \right\}\cup\\
    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}},\\ \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \endarray \right)} 
    \; \middle| \;
    \begin{array}{l}
      [\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss) \tand\\
      \forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
      \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal (\lrv \circ \xi)
    \end{array}
    \right\} \cup \\
    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
    \begin{array}{l}
      \begin{array}{r l l}
        \perm \not\in \{\rx,\rwx\} \wedge\\
        % \perm = \noperm & \Rightarrow & \npair{(\linear,\baddr,\eaddr)} \in \lrp(W) \wedge \\
        \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \stackReadCond{W} \wedge \\
        \perm \in \writeAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \stackWriteCond{W}
      \end{array}
    \end{array}
    \right\} \cup \\
%
    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
    \begin{array}{l}
      [b,e] \mathrel{\#} \ta \tand\\
      \begin{array}{r l l }
        % \perm = \noperm & \Rightarrow & \npair{(\lin,\baddr,\eaddr)} \in \lrp(W) \wedge \\
        \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \readCond{\lin,W} \wedge\\
        \perm \in \writeAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \writeCond{\lin,W} \wedge\\
        % we are excluding rwx pointers.
        \perm \neq \rwx \wedge \\
        % \perm = \rwx &\Rightarrow& \array[t]{l}\npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge \\
        %                            \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \endarray\\
        \perm = \rx &\Rightarrow& \array[t]{l}\npair{[\baddr,\eaddr]} \in \execCond{W} \wedge\\
        \npair{[\baddr,\eaddr]} \in \xReadCond{W} \wedge \\
                                  \lin = \normal \\ \endarray
      \end{array}
    \end{array}
    \right\}
  \end{array}
\]

\[
  \lrvtrusted[\square,\gc](W) =
  \begin{array}[t]{l}
    \lrv(W)\cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}},\\ \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \endarray \right)} 
    \; \middle| \;
    \begin{array}{l}
      \gc = (\ta,\stkb,\gsigrets,\gsigcloss)  \wedge \\
      \exists r \in \dom(\pwheap) \ldotp \\
      \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)} \wedge\\
      \quad \dom(\code) \subseteq \ta \wedge [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss) \wedge\\
      \quad \sigrets \subseteq \gsigrets \wedge \sigcloss \subseteq \gsigcloss
    \end{array}
    \right\} \cup \\
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\normal),\baddr,\eaddr,\aaddr)},\\ ((\perm,\normal),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
    \begin{array}{l}
      \perm \sqsubseteq \rx \wedge \\
      \gc = (\ta,\stkb,\gsigrets,\gsigcloss)  \wedge \\
      {} [\baddr,\eaddr] \subseteq \ta \wedge\\
      \npair{[\baddr,\eaddr]} \in \xReadCond[\square,\gc]{W} 
    \end{array}
    \right\}
  \end{array}
\]

@@

\begin{lemma}
  \label{lem:code-reg-and-mem-sat}
  If $\pwheap(r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$ and
  $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, then there exists $W''$
  and $W'$ such that $W = W' \oplus W''$ and
  $\npair[n']{(\code,\code)} \in H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; \gc \; \xi^{-1}(W')$ for all $n' < n$
\end{lemma}
\begin{proof}
  By definition of $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, we get that $W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$ and
  \begin{equation*}
    \npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in
    \lrheap(\pwheap)(W_{\var{heap}})
  \end{equation*}
  for some $\ms_S' \subseteq \ms_S$ and $\ms_{T,\var{heap}}\subseteq \ms_T$.

  By definition of $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in
  \lrheap(\pwheap)(W_{\var{heap}})$, we get an
  $R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg$ and $\exists R_W : \dom(\activeReg{\pwheap}) \fun \World$ such that
  \begin{itemize}
  \item $W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r)$ and
  \item $\npair[n']{R_\ms(r_{\var{code}})} \in \pwheap(r_{\var{code}}).H \; \xi^{-1}(R_W(r_{\var{code}}))$ for all $n' < n$
  \end{itemize}

@@

\begin{lemma}[non-linear words are pure]
  \label{lem:non-linear-pure}
  \hfill
  \begin{itemize}
  \item If $\npair{A} \in \readCond{\normal,W}$, then $\npair{A} \in \readCond{}(\normal, \purePart{W} )$.
  \item If $\npair{A} \in \writeCond{\normal,W}$, then $\npair{A} \in \writeCond{}(\normal,\purePart{W})$
  \item If $\npair{A} \in \execCond{W}$, then $\npair{A} \in \execCond{}(\purePart{W})$
  \item If $\npair{A} \in \xReadCond{W}$, then $\npair{A} \in \xReadCond{}(\purePart{W} )$.
  \item If $\npair{(w_1,w_2)} \in \lrv(W)$ and ($\nonLinear{w_1}$ or
    $\nonLinear{w_2}$), then
    $\npair{(w_1,w_2)} \in \lrv(\purePart{W})$.
  \end{itemize}
\end{lemma}
\begin{proof}
  Follows easily by inspecting the definitions of $\lrv$, $\readCond[]{}$, $\addressable{}$, $\writeCond[]{}$, $\execCond[]{}$ and $\xReadCond[]{}$ and using Lemma~\ref{lem:purePart-idempotent}.
\end{proof}

\begin{lemma}[permission-based conditions shrinkable]
  \label{lem:conds-shrinkable}
  \hfill
  \begin{itemize}
  \item If $\npair{A} \in \readCond{\lin,W}$  and $\emptyset \neq A' \subseteq A$, then
    $\npair{A'} \in \readCond{\lin,W}$.
  \item If $\npair{A} \in \writeCond{\lin,W}$  and $\emptyset \neq A' \subseteq A$, then
    $\npair{A'} \in \writeCond{\lin,W}$.
  \item If $\npair{A} \in \xReadCond{W}$  and $A' \subseteq A$, then
    $\npair{A'} \in \xReadCond{W}$.
  \item If $\npair{A} \in \execCond{W}$  and $\emptyset \neq A' \subseteq A$, then
    $\npair{A'} \in \execCond{W}$.
  \end{itemize}
\end{lemma}
\begin{proof}
  Follows easily from the definitions.
  In the case for $\execCond{}$ where $\lin = \linear$, we have a partition of the world into $W = \oplus_{a \in A} W_a(a)$ that we need to convert into a partition $W = \oplus_{a \in A'} W_a'(a)$.
  It suffices construct new $W_a'$ by taking $W_a$ but adding $\oplus_{a\in(A\setminus A')}$ to an $W_a(a)$ for an arbitrary $a \in A'$ to make this work.
\end{proof}

@@

   Take $R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg$ such that
    \begin{itemize}
    \item $R_\ms(r_\addr) = (\msdata|_{\{a\}}, \msdata|_{\{a\}})$.
    \item $R_\ms(r_\var{code}) = (\mspad\uplus\mscode,\mspad\uplus\mscode)$
    \end{itemize}
    Take $R_W' : \dom(\activeReg{\pwheap}) \fun \World$ such that
    \begin{itemize}
    \item $R_W'(r_\addr) = \purePart{W}[\var{heap}.r_{\addr'}\mapsto \stdreg{\{\aaddr\},\gc}{\spatialo}]_{\addr' \in A_\var{own}(\addr)}$ 
    \item $R_W'(r_\var{code}) = \purePart{W}$
    \end{itemize}
    Since then $W = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W'(r)$, we can use Lemma~\ref{lem:oplus-future-distr} to construct an $R_W$ with
    \begin{equation*}
      W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r)
    \end{equation*}
    and $R_W(r) \future R_W'(r)$ for all $r$.
    Finally, take $R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal}$ to satisfy
    \begin{itemize}
    \item $R_\var{seal}(r_\addr) = \emptyset$
    \item $R_\var{seal}(r_\var{code}) = \sigrets\uplus \sigcloss$.
    \end{itemize}

@@

  From $\npair{(\var{comp},\var{comp})} \in \lrcomp(W)$, we know that
  \begin{itemize}
  \item $\{(\_ \mapsto c_{\mathrm{main},c}), (\_ \mapsto c_{\mathrm{main},d})\}\subseteq \overline{\var{export}}$ and
  \item $\npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W)$ with
  \item  $\var{comp}_0 = (\mscode, \msdata, \overline{a_{\mathrm{import}}\mapsfrom s_{\mathrm{import}}}, \overline{s_{\mathrm{export}}\mapsto w_{\mathrm{export}}}, \sigrets, \sigcloss,A_\linear)$
  \end{itemize}
  From this, it follows that for all  $W' \future W$, if
  \begin{equation*}
    \overline{\npair[n']{(w_{\mathrm{import}},w_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}
  \end{equation*}
  and $\msdata' = \msdata{}[\overline{a_{\mathrm{import}} \mapsto w_{\mathrm{import}}}]$
  then
  \begin{equation*}
    \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W')
  \end{equation*}
  and
  \begin{equation*}
    \overline{\npair{(w_{\mathrm{export}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
  \end{equation*}
  Since $\overline{w_{\mathrm{import}}} = \emptyset$, the former holds vacuously, $\msdata' = \msdata$ and the latter two results follow for every $W' \future W$.
